### 2004-05-17 Documentation/part2.tex
*** empty log message ***

\section{A simple example for the AEDL}

@@

Eventually, we can write down the number of events per bin $i$ \index{Bin} and channel $c$ as
\begin{equation}
\label{eq:channel}
n_i^c=\int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' \quad
\frac{dn_{\beta}^{\text{IT}}}{dE'} (E').
\end{equation}
This means that one has to solve this integral
\begin{eqnarray}
\label{eq:events_bin}
n_i^c=N/L^2\,\int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' 
\quad \int\limits_0^\infty dE \,\, \Phi^c(E)\,
P^c(E)\,
\sigma^c(E)\,
R^c(E,E')\,
\epsilon^c(E')\,.
\end{eqnarray} 
Note that the events are binned according to their \emph{reconstructed} energy.

@@

% WW: HIER WEITER

@@

The definition of the energy resolution function is rather sophisticated in 
\GLOBES . In particular, since this aspect is rather time-critical, several
algorithms are implemented. The choice of the proper algorithm depends on the
experiment and the frequencies of the involved neutrino oscillations.

@@

@@

DESCRIBE PICTURE!

The origin and meaning of the energy resolution function 
$R^c(E,E')$ was already discussed in 
section~\ref{sec:channel} and  a mathematical definition
was given in equation~\ref{eq:e_res}. Here we deal with
how to tell \GLOBES\ which energy resolution function it should
use. This also naturally involves the issue how the actual convolution
of the differential event rate spectrum with $R_\beta^{\text{IT}}(E,E')$
is performed. First we will discuss the mathematical structure of
the problem and then the various algorithms and possibilities which
\GLOBES\ offers are introduced.

Before going into the details we define that the range from $E_\mathrm{min}$
to $E_\mathrm{max}$ is divided in a certain number of bins $N_\mathrm{bins}$.
These bins form a partition, which not necessarily is equidistant. The 
partition is any case completely defined by specifying $E_\mathrm{min}$
and $E_\mathrm{max}$ and the size of each bin  $\Delta E_i$. It is also
straightforward to compute the midpoint of each bin $E_i$. This information
has to be provided by the user. Equidistant bins are obtained by
\begin{quote}
{\tt
\$emin = 4.0\\
\$emax = 50.0\\
\$bins = 20
}
\end{quote}
Another possibility is to specify arbitrary bin sizes and to leave out
the number of bins itself, by
\begin{quote}
{\tt
\$emin = 4.0\\
\$emax = 50.0\\
\$binsize = \{  15.0 , 5.0 , 20.0, 6.0 \} 
}
\end{quote}

Instead of using directly equation~\ref{eq:e_res} we will use a different
definition concerning $\epsilon(E')$. In general  $\epsilon(E')$ has to be
determined by means of a Monte Carlo simulation of the experiment. This
usually involves anyhow a binning of the simulated events in the reconstructed
energy $E'$, therefore one more easily can obtain  $\hat\epsilon_i$ defined
by
\begin{equation}
\label{eq:post_smearing}
\hat\epsilon_i \cdot \int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' 
\quad R^c(E,E')\,.
\end{equation}
These  $\hat\epsilon_i$ are refered to as so called 
``post-smearing'' efficiencies and set within the corresponding {\tt channel}
environment.
The integration with respect to the reconstructed energy $E'$ can be
performed independently of the oscillation parameters and we define
the bin kernel for the $i$th bin $K_i^c$
\begin{equation}
\label{eq:kernel}
K_i^c(E):=\int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' 
\quad R^c(E,E')\,.
\end{equation}
With this definition equation~\ref{eq:events_bin} can be rewritten as
\begin{equation}
\label{eq:simple_int}
n_i^c=N/L^2 \int dE\quad  \underbrace{\Phi^c(E)\,
P^c(E)\,
\sigma^c(E)\,
K_i^c(E)\,}_{f(E)}. 
\end{equation}

There is no basic reason why one could not evaluate this integral directly
by the usual numerical methods. However it turns out that this may be rather
slow, therefore we introduce two different approximation schemes which have
a varying degree of accuracy and speed. In any case each scheme to evaluate this
integral involves a the evaluation of the integrand at fixed sampling points.
These sampling points can be either defined by the user or be computed by
\GLOBES, the choice between these possibilities also depends on the algorithm
which is used for the actual calculation, as explained in the following.
If the user chooses to set the sampling points the procedure is analogous
to the one setting the bins. For equidistant sampling points use
\begin{quote}
{\tt
\$simbins = 20\\
\$simtresh =          4.0\\
\$simbeam =         50.0
}
\end{quote}
The advantage in setting the sampling points in this way is that in case many
different energy resolution functions are used the user has the control
how the different sampling requirements are fulfilled. In case a user-defined
smear matrix is used this data is mandatory.

All energy resolution function are defined within an {\tt energy} environment
and can be refered to by {\tt \#name}.
\begin{quote}
  {\tt energy(\#name)<\\
\tb $\ldots$\\
>}
\end{quote}

ADJUST TO SECTION AND REPLACE:

In combination with the channels, is is possible to define energy 
dependent efficiencies and (non-beam) backgrounds. Efficiencies are multiplicative factors, whereas backgrounds are added to the event rates. These components can be introduced before or after the convolution in \eq~(\ref{eq:e_res}) is done. If they are introduced before, we call them
{\tt @pre\_smearing\_efficiencies} or {\tt @pre\_smearing\_background}. 
If they are introduced after, we call them {\tt @post\_smearing\_efficiencies} or {\tt @post\_smearing\_background}.
Note that pre-smearing components are always afunction of the incident neutrino energy $E$. Thus, there have to be as  many elements as there are sampling points (see \Sec~\ref{sec:energy}). In same cases, the number of sampling points is determined automatically by \GLOBES , and the number of elements in the  pre-smearing vectors has to be adjusted in the second run.
Examples for pre-smearing quantities are non-beam backgrounds, such as from geophysical neutrinos. The post-smearing components are always a function of the reconstructed neutrino energy $E'$, such as the post-smearing efficiencies $\epsilon_\beta^{\text{IT}}(E')$ in \eq~(\ref{eq:e_res}). Examples for post-smearing efficiencies are cuts and detection threshold functions. All post-smearing components have to have as 
many elements as there are energy bins. Efficiencies are multiplicative 
and their default value is $1$, whereas backgrounds are additive and their default value is $0$. Thus, a more elaborate channel can be defined as
\begin{quote}
{\tt channel(\#channel\_1)<\\
\tb @channel = \#flux : $+$: muon: muon: \#cross: \#energy\\
\tb @pre\_smearing\_background = \{1,2,3,4,5,6,7,8,9,10\}\\
\tb @post\_smearing\_efficiencies = \{0.1,0.2,0.3,0.4,0.5\}\\
>}
\end{quote}
This experiment uses $10$ sampling points and $5$ bins.

>>>>

\section{A simple example for AEDL}

@@

Eventually, we can write down the number of events per bin $i$ \index{Bin} and channel $c$ as
\begin{equation}
\label{eq:channel}
n_i^c=\int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' \quad
\frac{dn_{\beta}^{\text{IT}}}{dE'} (E') \,
\end{equation}
where $\Delta E_i$ is the bin size of the $i$th energy bin.
This means that one has to solve the integral
\begin{eqnarray}
\label{eq:events_bin}
n_i^c=N/L^2\,\int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' 
\quad \int\limits_0^\infty dE \,\, \Phi^c(E)\,
P^c(E)\,
\sigma^c(E)\,
R^c(E,E')\,
\epsilon^c(E')\,.
\end{eqnarray} 
Note that the events are binned according to their \emph{reconstructed} energy.

@@

@@

The definition and implementation of the energy resolution function is 
rather sophisticated in \GLOBES . In particular, since this aspect is 
 time-critical, several algorithms are implemented. The choice of the 
proper algorithm depends on the experiment and the frequencies of the
 involved neutrino oscillations.

@@

The energy resolution function $R^c(E,E')$ has been already introduced in 
\Sec~\ref{sec:channel}, where a definition
has been given in \eq~(\ref{eq:e_res}). Instead of using
 \eq~(\ref{eq:e_res}) directly, we apply a slightly different
definition of the post-smearing efficiencies $\epsilon(E')$. 
In general, $\epsilon(E')$ has to be
determined by means of a Monte Carlo simulation of the experiment. 
This usually involves a binning of the simulated events in the 
reconstructed energy $E'$. Therefore, one simplify \eq~(\ref{eq:events_bin}) by
\begin{equation}
\label{eq:post_smearing}
\int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' \, R^c(E,E') \, \epsilon^c(E') = \hat\epsilon_i^c \cdot \int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' 
\quad R^c(E,E')\,.
\end{equation}
Here the  $\hat\epsilon_i^c$ are the 
``post-smearing'' efficiencies, which will be set within the corresponding {\tt channel} environment (see below).
From \eq~(\ref{eq:events_bin}) it is obvious that the integration with respect to the reconstructed energy $E'$ can be
performed independently of the oscillation parameters. We define
the ``bin kernel'' $K_i^c$ for the $i$th bin as
\begin{equation}
\label{eq:kernel}
K_i^c(E) \equiv \int_{E_i-\Delta E_i/2}^{E_i+\Delta E_i/2} dE' 
\quad R^c(E,E')\,.
\end{equation}
With this definition, \eq~(\ref{eq:events_bin}) can be re-written as
\begin{equation}
\label{eq:simple_int}
n_i^c=N/L^2 \int\limits_0^\infty dE\quad  \underbrace{\Phi^c(E)\,
P^c(E)\,
\sigma^c(E)\,
K_i^c(E)\,}_{f(E)}. 
\end{equation}

There is no principle reason why one should not evaluate this integral directly by the usual numerical methods. However, it turns out that this 
is very slow in many cases. Therefore, we will introduce two different approximation schemes for different applications in the next two subsections.
In either case, the integrand in \eq~(\ref{eq:simple_int}) has to be evaluated at fixed ``sampling points''. These sampling points can be 
either defined by the user, or be computed by
\GLOBES . The choice between these options also depends on 
the algorithm which is used for the calculation, as it is explained in the following subsections.

@@

Before we come to the calculation algorithms, it is useful to understand
the general evaluation algorithm. As it is illustrated in \figu{energies}, 
\GLOBES\ uses several levels with respect to the energy ranges:
\begin{description}
\item[Sampling point level]
 This level is used internally to evaluate the integrand in \eq~(\ref{eq:simple_int}) at all sampling points. The energy scale is the actual incident neutrino energy $E$. Note that the sampling range should exceed the energy bin range by about the energy resolution error, since these sampling points need to be evaluated by the algorithm. However, this is only
necessary to avoid aliasing effects if the integrand in \eq~(\ref{eq:simple_int}) is large at the limits of the sampling range.
For a manual definition of the sampling points (depends on algorithm), use
\begin{quote}
{\tt
\$sampling\_points = 20\\
\$sampling\_min =          4.0\\
\$sampling\_max =         50.0
}
\end{quote}
for equidistant sampling points. Arbitrary sampling points can be specified analogously to the bins (see below).
\item[Bin level] This level is determined by the experiment. Note than energy
bin sizes much smaller than the energy resolution will not improve the results. The energy bin range and the number of energy bins do always have to be specified. For the case of large values of the integrand in \eq~(\ref{eq:simple_int}) at the energy range limits, it is recommended to exceed the analysis energy window by about the energy calibration error in order to avoid aliasing effects.

In order to define a range between $E_\mathrm{min}$
and $E_\mathrm{max}$ divided by a certain number of equidistant bins,
use
\begin{quote}
{\tt
\$emin = 4.0\\
\$emax = 50.0\\
\$bins = 20
}
\end{quote}
For arbitrary bins, use $E_\mathrm{min}$
and $E_\mathrm{max}$ and the size of each bin  $\Delta E_i$:
\begin{quote}
{\tt
\$emin = 4.0\\
\$emax = 50.0\\
\$binsize = \{  15.0 , 5.0 , 20.0, 6.0 \} 
}
\end{quote}
The number of bins will be automatically computed by \GLOBES . Note that the
bin sizes have to add up to the energy range {\tt \$emax}-{\tt \$emin}.
%
It is also
straightforward to compute the midpoint of each bin $E_i$. This information
has to be provided by the user ???? HOW ????. 
\item[Analysis level]
On the analysis level, an energy window can be defined within each rule. For
details, see next chapter.
\end{description}
In general, the energy smearing happens between the sampling point and bin levels, which means that the energy smearing matrix will have {\tt \$sampling\_points} columns and {\tt \$bins} rows.

As illustrated in the figure, an interesting feature in combination with the
channels are pre- and post-smearing effects. Pre-smearing effects are taken into account on the sampling point level, and post-smearing effects on the bin level. Examples for these effects are energy 
dependent efficiencies and (non-beam) backgrounds. Efficiencies are multiplicative factors, whereas backgrounds are added to the event rates. These components can be introduced before or after the convolution in \eq~(\ref{eq:post_smearing}) is done. If they are introduced before, 
we call them
{\tt @pre\_smearing\_efficiencies} or {\tt @pre\_smearing\_background}. 
If they are introduced after, we call them {\tt @post\_smearing\_efficiencies} or {\tt @post\_smearing\_background}.
Note that pre-smearing components are always afunction of the incident neutrino energy $E$. Thus, there have to be as  many elements as there are sampling points (see \Sec~\ref{sec:energy}). In same cases, the number of sampling points is determined automatically by \GLOBES , and the number of elements in the  pre-smearing vectors has to be adjusted in the second run.
Examples for pre-smearing quantities are non-beam backgrounds, such as from geophysical neutrinos. The post-smearing components are always a function of the reconstructed neutrino energy $E'$, such as the post-smearing efficiencies $\epsilon_\beta^{\text{IT}}(E')$ in \eq~(\ref{eq:e_res}). Examples for post-smearing efficiencies are cuts and detection threshold functions. All post-smearing components have to have as 
many elements as there are energy bins. Efficiencies are multiplicative 
and their default value is $1$, whereas backgrounds are additive and their default value is $0$. Thus, a more elaborate channel can be defined as
\begin{quote}
{\tt channel(\#channel\_1)<\\
\tb @channel = \#flux : $+$: muon: muon: \#cross: \#energy\\
\tb @pre\_smearing\_background = \{1,2,3,4,5,6,7,8,9,10\}\\
\tb @post\_smearing\_efficiencies = \{0.1,0.2,0.3,0.4,0.5\}\\
>}
\end{quote}
This experiment uses $10$ sampling points and $5$ bins.

In the following subsections we will define the energy resolution function.
All energy resolution functions are defined within an {\tt energy} environment and can be refered to by {\tt \#name}.
\begin{quote}
  {\tt energy(\#name)<\\
\tb $\ldots$\\
>}
\end{quote}

%% WW: CONTINUE HERE
